---
title: "Project"
author: "Anish"
date: "1/22/2018"
output: html_document
---

# PRACTICAL MACHINE LEARNING PROJECT 

## OBJECTIVE  

The objective of this project is to make use of the data from accelerometers on the belt, forearm, arm and dumbell of 6 participants to predict the manner in which they did the exercise

## DATA PREPROCESSING  

Loading required packages

```{r}
library(caret)
library(rpart)
```

Reading the data from the training and testing csv files

```{r}
training = read.csv("/Users/anish-zt70/Downloads/SONGS/pml-training.csv")
test = read.csv("/Users/anish-zt70/Downloads/SONGS/pml-testing.csv")
```

Setting the seed , so that the experiment can be reproduced

```{r}
set.seed(10)
```

Creating the training and validation data 

```{r}
inTrain <- createDataPartition(y=training$classe, p=0.7, list=FALSE)
trainData <- training[inTrain, ]
crossValidData <- training[-inTrain, ]
```

### PreProcessing

Removing the columns that doesn't provide much information for prediction such as user_name, raw_timestamp_part1 etc

```{r}
trainData <- trainData[,-c(1:7)]
```

Removing the columns that have near zero variance since it doesn't have much statistical importance

```{r}
nzVar <- nearZeroVar(trainData)
trainData <- trainData[, -nzVar]
```

Removing columns which have "na" values in it 

```{r}
naColumns <- colSums(is.na(trainData)) == 0
trainData <- trainData[, naColumns]
```

```{r}
head(trainData)
```

Obtaining the column information from training set so that they can be used in the cross validation data set also

```{r}
trainingColumns <- colnames(trainData)
crossValidData <- crossValidData[, trainingColumns]
```

### GBM Model

Tried using the gradient boosting method by adjusting the default parameters to predict the data in a minimal time

```{r}
trControl <- trainControl(method = "cv", number = 2)
```

Training the GBM Model

```{r}
fitGBM <- train(classe ~ ., data=trainData, method = "gbm", trControl=trControl)
fitGBM$finalModel
```

Using the model to predict the outputs

```{r}
predGBM <- predict(fitGBM, crossValidData)
confusionMatrix(crossValidData$classe, predGBM)$overall[1]
```

The accuracy obtained using the above method is 
```{r}
confusionMatrix(crossValidData$classe, predGBM)$overall[1]
```

### Random Forest Model

Replacing random forest instead of GBM to see the predictions in the above mode

```{r}
fitRf <- train(classe ~ ., data=trainData, method = "rf", trControl=trControl)
fitRf$finalModel
predRf <- predict(fitRf, crossValidData)
confusionMatrix(crossValidData$classe, predRf)
```

The accuracy obtained using the Random Forest method is 

```{r}
confusionMatrix(crossValidData$classe, predRf)$overall[1]
```

### Storing the predicted Values

```{r}
predRf <- as.character(predRf)
writeFiles <- function(x) {
  n <- length(x)
  for(i in 1:n) {
    filename <- paste0("problem_id_", i, ".txt")
    write.table(x[i], file=filename, quote=F, row.names=F, col.names=F)
  }
}
```

```{r}
writeFiles(predRf)
```